---
title: "DR2 VWP data manipulation"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(gazer)
```

### Load data:

This step involves loading 41 of the participant data files and then converting all columns to character values. This was a necessary step to autoload all data. I will change data types as it becomes necessary below.

```{r}
# folder_path = 'all_data/updated_xls_format'
# all_files = list.files(path = folder_path, pattern = '\\.xlsx$|\\.xls$', full.names = TRUE)
# 
# df_list = list()
# 
# for (file in all_files) {
#   temp_df <- readxl::read_excel(file)
#   temp_df$file_name <- file  # Adding the file_name column
#   
#   # Convert all columns to character
#   temp_df <- mutate_all(temp_df, as.character)
#   
#   df_list <- append(df_list, list(temp_df))
# }
# 
# # Concatenate all dataframes in the list into a single dataframe
# df = bind_rows(df_list)
# head(df)

```

Next step: extract participant and session names:

```{r}
df = df %>%
  mutate(participant = str_extract(file_name, "(?<=updated_xls_format/)\\d+"),
         session = str_extract(file_name, "(?<=_)\\d+"))

```

Remove practice conditions

```{r}
df = df %>%
  filter(condition != 'practice' )
```

Combine left and right eye data

```{r}
df <- df %>%
  mutate(aoi1 = ifelse(RIGHT_IA_1_SAMPLE_COUNT != '.', RIGHT_IA_1_SAMPLE_COUNT, LEFT_IA_1_SAMPLE_COUNT),
         aoi2 = ifelse(RIGHT_IA_2_SAMPLE_COUNT != '.', RIGHT_IA_2_SAMPLE_COUNT, LEFT_IA_2_SAMPLE_COUNT),
         aoi3 = ifelse(RIGHT_IA_3_SAMPLE_COUNT != '.', RIGHT_IA_3_SAMPLE_COUNT, LEFT_IA_3_SAMPLE_COUNT),
         aoi4 = ifelse(RIGHT_IA_4_SAMPLE_COUNT != '.', RIGHT_IA_4_SAMPLE_COUNT, LEFT_IA_4_SAMPLE_COUNT))
```

```{r}
df = df %>%
  mutate(off_target = as.numeric(aoi1) + as.numeric(aoi2) + as.numeric(aoi3) + as.numeric(aoi4)) 


df = df %>%
  filter(off_target > 0)
```

Create count variables

```{r}
df2 = df %>%
    pivot_longer(cols = aoi1:aoi4,
               names_to = "fixation_location",
               values_to = "count")

df2$fixation_location = as.numeric(substr(df2$fixation_location, 4,4))
```

```{r}
df3 = df2 %>%
  mutate(fixation_name = 
           ifelse(fixation_location == target_location, 'Target',
           ifelse(fixation_location == phonemic_location, "Phonemic",
           ifelse(fixation_location == semantic_location, "Semantic",
           ifelse(fixation_location == unrelated_location, "unrelated", 'Error')      
                  )))) %>%
  mutate(count = as.numeric(count))
```

```{r}
df3 = df3 %>%
  mutate(count_round = ifelse(as.numeric(count) > 12.1, 1, 0))
```

```{r}
df3 %>% 
  na.omit() %>%
  ggplot(aes(x = BIN_INDEX, y = count_round)) +
  geom_bar(stat = 'summary') +
  facet_wrap(~fixation_name)

```

```{r}
test = df3 %>%
  filter(fixation_name == 'unrelated')

df3 %>% group_by(fixation_name) %>%
  summarize(sum(count))
```

```{}
```

Try to get proportions again - Mirman groups by condition here, which for us would mean treated,untreated, unrelated.

```{r}
df4 = df3 %>%
  filter(RESPONSE == 'target') %>%
  filter(BIN_START_TIME < 3500) %>%
  group_by(participant, session, fixation_name) %>%
  mutate(nTrials = length(unique(TRIAL_INDEX))) %>% 
  ungroup() %>%
  # calculate number of fixations 
  group_by(participant, session, fixation_name, BIN_INDEX) %>%
  summarize(sumFix = sum(count_round), nTrials = unique(nTrials),
            meanFix = sum(count_round)/unique(nTrials)) %>%
  na.omit(meanFix)
```

```{r}
ggplot(df4, aes(as.numeric(BIN_INDEX), meanFix, colour = fixation_name)) +
  #facet_wrap(~ session) +
  stat_summary(fun = 'mean', geom = "line")
```
