---
title: "Data aggregation investigation"
author: "AMS"
date: "3/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(readr)
#working directory package
library(here)
#to load data
library(readxl)
library(readr)
#to manipulate data, dplyr, purr, etc and to plot w ggplot2
library(tidyverse)
library(janitor)
library(reshape2)
#Plot tree structure
library(data.tree)
#Frequentist multilevel models
library(lme4)
#Bayesian multilevel models
library(brms)
#pretty plots
library(sjPlot)
library(stargazer)
library(factoextra)
###necessary for downlaoding and install factoextra, nice package for plotting and summarizing opca
#if(!require(devtools)) install.packages("devtools")
#devtools::install_github("kassambara/factoextra")

```

```{r, include=FALSE}
#load data
here()
dr1 = read_csv(here("DR1PCADATA.csv"))
dr2 = read_excel(here("VWP_PreTx_assessments.xlsx"))
data = clean_names(dr1)
```

### fix palpa 49 outlier, assuming it is a typo

```{r}
data = data %>%
  mutate(ifelse(palpa49_audsyn_hi == 96.6700, 0.96700, palpa49_audsyn_hi))

summary(data$palpa49_audsyn_hi)
```

```{r}
#create data set for pca
data2 = data.frame(
  #sweight = data$s_weight,
  #pweight = data$p_weight,
  ppt = data$ppt,
  #p_nwREP_count = data$palpa8_nword,
  aud_rhyme_OA = data$palpa15_a_overall,
  written_rhyme_OA = data$palpa15_w_overall,
  aud_syn_OA = data$palpa49_audsyn_overall,
  written_syn = data$palpa_50_written_synonym_overall,
  CAT = data$pre_mean,
  participant = as.factor(seq(1:nrow(data))))

stargazer(data.frame(data2), type = "text", title = "Table 1: DR1 Sample", digits = 2, align = TRUE)


dr2namechange = data.frame(ppt = dr2$ppt/52,
                           aud_rhyme_OA = dr2$palpa15_aud_rhyme / 60,
                           written_rhyme_OA = dr2$palpa15_written_rhyme / 60,
                           aud_syn_OA = dr2$palpa49_aud_syn / 60, 
                           written_syn = dr2$palpa50_written_syn/60,
                           CAT = dr2$cat_mmt_1,
                           participant = dr2$participant)


stargazer(data.frame(dr2namechange), type = "text", title = "Table 2: DR2 Sample", digits = 2, align = TRUE)

pcadata = rbind(data2, dr2namechange)

stargazer(data.frame(pcadata), type = "text", title = "Table 3: DR1&2 Sample", digits = 2, align = TRUE)
#following suggestions from gary dell, take square root of s and p for analysis.
# data3 = data2 %>%
#   #mutate(sweight_sqr =  sqrt(sweight)) %>%
#   #mutate(pweight_sqr = sqrt(pweight)) %>%
#   #remove old estimates of s and p
#     select(c(-sweight, -pweight)) 
```

## correlation matrix

```{r}
correlation = round(cor(pcadata[,1:6]),2)
stargazer(correlation, summary = FALSE, type = "text", title = "Table 4: DR1 and 2 correlation matrix", align = TRUE, digits = 2)
```

#Formating data for PCA - Need to create equal variance, mean centering

```{r}
#df_center = data.frame(sapply(pcadata, function(x) scale(x, scale = FALSE)))
df_center = data.frame(ppt = scale(pcadata$ppt, scale = TRUE ),
                       aud_rhyme = scale(pcadata$aud_rhyme_OA, scale = TRUE),
                       written_rhyme = scale(pcadata$written_rhyme_OA, scale = TRUE),
                       aud_syn = scale(pcadata$aud_syn_OA, scale = TRUE),
                       written_syn = scale(pcadata$written_syn, scale = TRUE)
                       )

stargazer(data.frame(pcadata), type = "text", title = "Table 5: Raw variables (should have unequal variances)")
stargazer(data.frame(df_center), type = "text", title = "Table 6: Mean Centered Variables (should have equal variance)")
```

### Check distributions

```{r}
#inspect distributions

df_center %>%
  mutate(Participant = seq(1: nrow(df_center))) %>%
  gather(Assessment, Estimate, ppt:written_syn) %>%
  ggplot(aes(x = Estimate)) +
  geom_histogram(color = "steelblue", size = 1, bins = 20) +
  labs(title = "Figure 2: Distribution of Dose Response 1 and 2 variables mean centered and z-scored") +
  facet_wrap(~ Assessment)
```

### need to z score to try and take care of these wide rangers of numbers

### Creating df3 to standardize CAT scores and fix annoying naming conventions

-   not using for now, but do you want to include any assessments from the CAT?

```{r}
#correlation matrix
cormat = round(cor(df_center), 2)
#rearrange data for heatmap looks like a 3 column correlation matrix
melted_cormat <- melt(cormat)
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
upper_tri <- get_upper_tri(cormat)
upper_tri
# Melt the correlation matrix
library(reshape2)
melted_cormat <- melt(upper_tri, na.rm = TRUE)

reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}
# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1)) +
  geom_text(aes(Var2, Var1, label = value), color = "black", size =4)  +
  ggtitle("Figure 3: DR 1 and 2 IVs, mean centered and z-scored")
# Print the heatmap
print(ggheatmap)
```

### correlations across variables are obviously present, assessing ability to use pca to create aggregate variables. Correlations amongst phonological variables are not obvious.

```{r}
#covmatrix = cov.wt(df_center)
#cm = data.frame(covmatrix$cov)
#pca_out = princomp(~ ppt + aud_rhyme + written_rhyme + aud_syn + written_syn, covmatrix  = cm, scores = T)
```

#Create df_center with all variables from dr2

```{r}
dr2_vars = data.frame(ppt = dr2$ppt/52,
                           aud_rhyme_OA = dr2$palpa15_aud_rhyme / 60,
                           written_rhyme_OA = dr2$palpa15_written_rhyme / 60,
                           aud_syn_OA = dr2$palpa49_aud_syn / 60, 
                           written_syn = dr2$palpa50_written_syn/60,
                           CAT = dr2$cat_mmt_1,
                           Nonword_rep = dr2$palpa8_nwor/30,
                           participant = dr2$participant)

dr2_center = data.frame(ppt = scale(dr2_vars$ppt, scale = TRUE ),
                       aud_rhyme = scale(dr2_vars$aud_rhyme_OA, scale = TRUE),
                       written_rhyme = scale(dr2_vars$written_rhyme_OA, scale = TRUE),
                       aud_syn = scale(dr2_vars$aud_syn_OA, scale = TRUE),
                       written_syn = scale(dr2_vars$written_syn, scale = TRUE),
                       CAT = scale(dr2_vars$CAT, scale = TRUE),
                       Nonword_rep = scale(dr2_vars$Nonword_rep, scale = TRUE),
                       Participant = dr2_vars$participant)

write.csv(dr2_center, "dr2_IVs.csv")
stargazer(data.frame(dr2_center), type = "text", title = "Table 6: DR2 all vars, mean centered and z-scaled", summary = FALSE, digits = 2, align = TRUE, rownames = TRUE)

#correlation matrix
cormat = round(cor(dr2_center[,1:7]), 2)
#rearrange data for heatmap looks like a 3 column correlation matrix
melted_cormat <- melt(cormat)
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
upper_tri <- get_upper_tri(cormat)
upper_tri
# Melt the correlation matrix
library(reshape2)
melted_cormat <- melt(upper_tri, na.rm = TRUE)

reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}
# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1)) +
  geom_text(aes(Var2, Var1, label = value), color = "black", size =4) +
  ggtitle("DR2 IVs mean centered and z-scored")
# Print the heatmap
print(ggheatmap)

write_csv(dr2_center, "dr2_center_DVs.csv")
```

### PCA results are weird, not going to try and interpret righ tnow. Recommend repeating once palpa 8 has been added to dataset

```{r}
#summary(pca_out)
```

```{r}
#pca_out$center
```

```{r}
#screeplot(pca_out)
```

### Packages:

```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
