---
title: "you down with GROUPED mpt"
author: "AMS"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Load required libraries
```{r}
#to load data
library(readxl)
library(readr)
#to manipulate data, dplyr, purr, etc and to plot w ggplot2
library(tidyverse)
#Plot tree structure
library(data.tree)
#Frequentist multilevel models
library(lme4)
#Bayesian multilevel models
library(brms)
#pretty plots
library(sjPlot)
```

```{r}
my_files = list.files("~/GitHub/DR2/Specific aim 2/Academy of Aphasia/data")

#Further arguments to read.csv can be passed in ...
all_csv <- lapply(my_files, read.csv)

#Set the name of each list element to its
# respective file name. Note full.names = FALSE to
# get only the file names, not the full path.
names(all_csv) <- gsub(".csv","",
                       list.files("~/GitHub/DR2/Specific aim 2/Academy of Aphasia/data",full.names = FALSE),
                       fixed = TRUE)

```

#manipulate data
```{r}
#remove practice
library(tidyverse)
list = all_csv %>%
  map(~filter(., condition != "practice"))
list_check = all_csv %>%
  map(~filter(., condition != "practice"))

```
#### Extract data sets that do not use right eye
- p11_02 uses left eye
- p13_01 uses right eye bus has no data in the interest area variables

```{r}
#create p11_2 df to manipulate and return to later.
p11_2 = list[['p11_2']]
#use pluck to remove p11_02 and 13_01 from list
pluck(list,  "p11_2") = NULL
pluck(list,  "p13_01") = NULL
```
#### Sanity check for above, list should now have two elements less that 



```{r}
list2_check = list
list2 = list %>%
  map(~rename(., aoi1 = RIGHT_IA_1_SAMPLE_COUNT)) %>%
  map(~rename(., aoi2 = RIGHT_IA_2_SAMPLE_COUNT)) %>%
  map(~rename(., aoi3 = RIGHT_IA_3_SAMPLE_COUNT)) %>%
  map(~rename(., aoi4 = RIGHT_IA_4_SAMPLE_COUNT))
```

#### Sanity check: aoi1-4 in list 2 should be equal to the right_IA_1... in list2_check.  checking a couple
```{r}
#participant 2 01
setequal(list2["p02_01"]$p02_01$aoi1, list2_check["p02_01"]$p02_01$RIGHT_IA_1_SAMPLE_COUNT)
setequal(list2["p02_01"]$p02_01$aoi2, list2_check["p02_01"]$p02_01$RIGHT_IA_2_SAMPLE_COUNT)
setequal(list2["p02_01"]$p02_01$aoi3, list2_check["p02_01"]$p02_01$RIGHT_IA_3_SAMPLE_COUNT)
setequal(list2["p02_01"]$p02_01$aoi4, list2_check["p02_01"]$p02_01$RIGHT_IA_4_SAMPLE_COUNT)

#participant 2 02 
setequal(list2["p02_02"]$p02_02$aoi1, list2_check["p02_02"]$p02_02$RIGHT_IA_1_SAMPLE_COUNT)
setequal(list2["p02_02"]$p02_02$aoi2, list2_check["p02_02"]$p02_02$RIGHT_IA_2_SAMPLE_COUNT)
setequal(list2["p02_02"]$p02_02$aoi3, list2_check["p02_02"]$p02_02$RIGHT_IA_3_SAMPLE_COUNT)
setequal(list2["p02_02"]$p02_02$aoi4, list2_check["p02_02"]$p02_02$RIGHT_IA_4_SAMPLE_COUNT)

#participant 13 01 
setequal(list2["p13_01"]$p13_01$aoi1, list2_check["p13_01"]$p13_01$RIGHT_IA_1_SAMPLE_COUNT)
setequal(list2["p13_01"]$p13_01$aoi2, list2_check["p13_01"]$p13_01$RIGHT_IA_2_SAMPLE_COUNT)
setequal(list2["p13_01"]$p13_01$aoi3, list2_check["p13_01"]$p13_01$RIGHT_IA_3_SAMPLE_COUNT)
setequal(list2["p13_01"]$p13_01$aoi4, list2_check["p13_01"]$p13_01$RIGHT_IA_4_SAMPLE_COUNT)

#participant 17 01
setequal(list2["p17_01"]$p17_01$aoi1, list2_check["p17_01"]$p17_01$RIGHT_IA_1_SAMPLE_COUNT)
setequal(list2["p17_01"]$p17_01$aoi2, list2_check["p17_01"]$p17_01$RIGHT_IA_2_SAMPLE_COUNT)
setequal(list2["p17_01"]$p17_01$aoi3, list2_check["p17_01"]$p17_01$RIGHT_IA_3_SAMPLE_COUNT)
setequal(list2["p17_01"]$p17_01$aoi4, list2_check["p17_01"]$p17_01$RIGHT_IA_4_SAMPLE_COUNT)
```


#### Following schmit et al, change all values below 5ms to 0 and above to 10 (pg 14, table expalnation.

```{r}
list3_check = list2
list3 = list2 %>%
  map(~mutate(., aoi1 = ifelse(aoi1 > 5, 10, 0))) %>%
  map(~mutate(., aoi2 = ifelse(aoi2 > 5, 10, 0))) %>%
  map(~mutate(., aoi3 = ifelse(aoi3 > 5, 10, 0))) %>%
  map(~mutate(., aoi4 = ifelse(aoi4 > 5, 10, 0)))
```

#### create a variable that shows us where the participant was looking based on numerical values. 
- logic, if aoi# is greater than zero, the participant is fixating on it. One aoi value for each time point in ms. 
```{r}
list4_check = list3
list4 = list3 %>%
  map(~mutate(., aoi_all = ifelse(aoi1 > 0, 1, #if aoi1 is greater than 0, label aoi_all 1, if not check below
                     ifelse(aoi2 > 0, 2, #if aoi2 is greater than 0, label 2 etc
                            ifelse(aoi3 > 0, 3,
                                   ifelse(aoi4 > 0, 4, 0)))))) 
```

#### Fill 0 values with the next number observed in the aoi_all variable
```{r}
list4 = list4 %>%
  map(~mutate(.,  na_if(aoi_all, 0)) %>% 
  fill(aoi_all, .direction = "up"))
```

#### the next step is to create a single column that shows what the participant is looking at by cominging the 
- logic: compare where the participant is looking at with variable aoi_all to the location of each interest area. *note, some participants did not have an distractor/target location so I had to create this variable based on the sum of the values of the other 3 location types. 
  -eg. there are 4 possible locations, therefore the sum of the known three can be used to find the location of the other. Consider the locations below. 
  $$s1+u2+p3 = 6$$ therefore, c = 4
  Repeat for all other locations
  $$s1 + u3 + p4 = 8$$ therefore, c = 2
  $$ s2 + u3 + p4 = 9$$ therefore, c = 1

```{r}
list4$p02_01$target_location = ifelse(list4$p02_01$semantic_location + list4$p02_01$phonemic_location + list4$p02_01$unrelated_location == 6, 4,
                                      ifelse(list4$p02_01$semantic_location + list4$p02_01$phonemic_location + list4$p02_01$unrelated_location == 8, 2, 1))
```

#### same logic for targep10_1, who is missing unrelated location
```{r}
list4$p10_1$unrelated_location = ifelse(list4$p10_1$semantic_location + list4$p10_1$phonemic_location + list4$p10_1$target_location == 6, 4,
                                      ifelse(list4$p10_1$semantic_location + list4$p10_1$phonemic_location + list4$p10_1$target_location == 8, 2, 1))

```

"target_location" variable. Therefore, using the last ifelse statement to label if participant is looking at target location. Only label y_5 = 1, if no other condition is met. 
```{r}

#create single column showing what aoi participant was looking at
list4 = list4 %>%
  map(~mutate(., y_5 = ifelse(aoi_all == unrelated_location, 4,
                 ifelse(aoi_all == phonemic_location, 3,
                        ifelse(aoi_all == semantic_location, 2, 1)))))
```

#### combine data for people with right eye fixation

```{r}
#first create  dummy variables for session number,  eye, participant #
#session number 
list4$p02_01$Session = rep(1)
list4$p02_02$Session = rep(2)
list4$p10_1$Session = rep(1)
list4$p10_2$Session = rep(2)
list4$p11_1$Session = rep(1)
list4$p13_02$Session = rep(2)
list4$p17_01$Session = rep(1)
list4$p17_02$Session = rep(2)

#eye
list4$p02_01$eye = rep(1)
list4$p02_02$eye = rep(1)
list4$p10_1$eye = rep(1)
list4$p10_2$eye = rep(1)
list4$p11_1$eye = rep(1)
list4$p13_02$eye = rep(1)
list4$p17_01$eye = rep(1)
list4$p17_02$eye = rep(1)

#particpant number
list4$p02_01$id =rep(02)
list4$p02_02$id =rep(02)
list4$p10_1$id =rep(10)
list4$p10_2$id =rep(10)
list4$p11_1$id =rep(11)
list4$p13_02$id =rep(13)
list4$p17_01$id =rep(17)
list4$p17_02$id =rep(17)
```
#### clean up targetword variable for anyone that doesnt have it

```{r}

list4$p10_1$targetword = list4$p10_1$target
```

```{r}
data_re = list4 %>%
  map(~mutate(., Target_actual = gsub("_.*","",target))) %>%
  map(~select(., TRIAL_INDEX, BIN_START_TIME, Target_actual, y_5, Session, eye, id)) %>%
  list.rbind()
```

#### OMG return back to left eye. 
```{r}
df2 = p11_2 %>%
  rename(aoi1 = LEFT_IA_1_SAMPLE_COUNT) %>%
  rename(aoi2 = LEFT_IA_2_SAMPLE_COUNT) %>%
  rename(aoi3 = LEFT_IA_3_SAMPLE_COUNT) %>%
  rename(aoi4 = LEFT_IA_4_SAMPLE_COUNT)

#Following schmit et al, change all values below 5ms to 0 and above to 10 (pg 14, table expalnation.
df2$aoi1 = ifelse(df2$aoi1 < 5, 10, 0)
df2$aoi2 = ifelse(df2$aoi2 < 5, 10, 0)
df2$aoi3 = ifelse(df2$aoi3 < 5, 10, 0)
df2$aoi4 = ifelse(df2$aoi4  < 5, 10, 0)

#create single column showing what aoi participant was looking at
df2$aoi_all = ifelse(df2$aoi1 > 0, 1,
                     ifelse(df2$aoi2 > 0, 2, 
                            ifelse(df2$aoi3 > 0, 3,
                                   ifelse(df2$aoi4 > 0,4, 0))))

df2$y_5 = ifelse(df2$aoi_all == df2$unrelated_location, 2,
                 ifelse(df2$aoi_all == df2$phonemic_location, 4,
                        ifelse(df2$aoi_all == df2$semantic_location, 3,
                               ifelse(df2$aoi_all == df2$target_location, 1, 0)))) 
# yy.list is the hierachical structure of available steps in the tree model. 

df2 = df2 %>%
  mutate(y =  na_if(y_5, 0)) %>% 
  fill(y, .direction = "up")

df2.5 = data.frame(  
  TRIAL_INDEX = df2$TRIAL_INDEX,
  BIN_START_TIME = df2$BIN_START_TIME,
  Target_actual = df2$targetword,
  y_5 = df2$y_5,
  Session = rep(2),
  eye = rep(2),
  id = rep(11)
)
```

#### Combine left and right eyes

```{r}
df3 = rbind(df2.5, data_re) %>% 
  rename(y =y_5)
df3_check = df3
```

#### Triplicate data for x3 nodes
```{r}
number.of.nodes <- 3
for (node in 1:number.of.nodes){
  data.copy <- df3
  data.copy$node <- rep(node, nrow(df3))
  if (node==1){data.node <- data.copy} else {data.node <- rbind(data.node,data.copy)}
}
```

```{r}
data.node$yy = ifelse(data.node$node== 1 & data.node$y < 4, 1,  #node 1 and correct (sem, unrelated, correct) 
                      ifelse(data.node$node== 1 & data.node$y == 4, 0, #node 1 and phon distractor
                             ifelse(data.node$node== 2 & data.node$y == 4, NA, #node 2 and phonological comp
                             ifelse(data.node$node== 2 & data.node$y == 3, 0, #node 2 and phonological comp
                             ifelse(data.node$node== 2 & data.node$y < 3, 1,  #node 2 and unrelated or correct
                             ifelse(data.node$node== 3 & data.node$y > 2, NA, #node 3 and y = phonological or semantic competitor
                             ifelse(data.node$node== 3 & data.node$y == 2, 0, 1 #node 3 and unrelated competitor, if not unrelated then correct
                             )))))))
unique(data.node$yy)

```

```{r}
if (length(which(data.node$node==2 & data.node$y==4))>0){
  data.node <- data.node[-which(data.node$node==2 & data.node$y==4),]
}
#Remove y = 3 or 4 for node 3
if (length(which(data.node$node==3 & data.node$y>2.1))>0){
  data.node <- data.node[-which(data.node$node==3 & data.node$y>2.1),]
}

#Test, node 2 has no y = 4, true
test_node_2 = data.node %>%
  filter(node == 2) %>%
  filter(y ==4)
nrow(test_node_2)
#Test, node 3 has no y = 3 or 4, true
test_node_3 = data.node %>%
  filter(node == 3) %>%
  filter(y > 2.1)
nrow(test_node_3)
```
#### ## Create variables node.1, node.2, and node.3
- node.1: if the participant is looking at phonological comp 0, anything else 1.
- node.2 At node 2, if the participant is looking at semantic 0, anything else (note phonological no longer in dataset for node = 2)
-node.3 At node 3, if the participant is looking at unrelated 0, and correct = 1 (note phonological and semantic competitors no longer in dataset for node = 2)
```{r}
#At node 1, if the participant is looking at unrelated comeptitor 0 anything else down the tree, 1
data.node$node.1 = ifelse(data.node$y == 4, 0, 1)
#At node 2, if the participant is looking at phonological, 0 anything else 1
data.node$node.2 = ifelse(data.node$y > 3, 0, 1)
#At node 3, if the participant is looking at correct, 1 semantic 0
data.node$node.3 = ifelse(data.node$y == 2, 0, 1)
```


## Create variable t and c
-t: Whether the person is looking at the target. t=1 if y=1, and t=0 otherwise.
-c: Whether the person is looking at the competitor. c=1 if y=2, and c=0 otherwise.



```{r}
node1_df = data.node %>%
  filter(node == 1) %>%
  rename(item = Target_actual) %>%
  mutate(t = ifelse(node1_df$y == 1, 1,0)) %>%
  mutate(c = ifelse(node1_df$y !=1,1,0))

node2_df = data.node %>%
  filter(node == 2) %>%
  rename(item = Target_actual) %>%
  mutate(t = ifelse(node2_df$y == 1, 1,0)) %>%
  mutate(c = ifelse(node2_df$y !=1,1,0))

node3_df = data.node %>%
  filter(node == 3)%>%
  rename(item = Target_actual) %>%
  mutate(t = ifelse(node3_df$y == 1, 1,0)) %>%
  mutate(c = ifelse(node3_df$y !=1,1,0))

data.node = rbind(node1_df, node2_df, node3_df)

#SANITY CHECK: ARE THERE NOW 3 NODES AGAIN?
unique(data.node$node)
```
#### Create lag variables: t.lag, c.lag, yy.lag
- t.lag: The value of t at the previous time point. Undefined for the first time point.
- c.lag: The value of c at the previous time point. Undefined for the first time point.
- yy.lag: The value of yy at the previous time point. Undefined for the first time point.
```{r}
data.node = data.node %>%
  arrange(node) %>%
  group_by(node, id, item) %>%
  mutate(t.lag = lag(t)) %>%
  mutate(c.lag = lag(c)) %>%
  mutate(yy.lag = lag(yy))%>%
  group_by() %>%
  mutate(time.coded.center = scale(BIN_START_TIME)) %>%
  filter(!is.na(t.lag))

#SANITY CHECK: Check to make sure t.lag, c.lag, nor yy.lag include NA 
summary(data.node)
```
## Declare variable structure for analyses:
```{r}
# create correct variable structures 
data.node$yy.lag <- as.numeric(data.node$yy.lag)
data.node$t.lag <- as.numeric(data.node$t.lag)
data.node$c.lag <- as.numeric(data.node$c.lag)
#data.node$time.coded <- as.numeric(data.node$time.coded)
data.node$time.coded.centered <- as.numeric(data.node$time.coded.center)
data.node$item <- as.factor(data.node$item)
data.node$person <- as.factor(data.node$id)
data.node$trial <- as.factor(data.node$TRIAL_INDEX)
data.node$node <- as.factor(data.node$node)
data.node$node.1 <- as.factor(data.node$node.1)
data.node$node.2 <- as.factor(data.node$node.2)
data.node$node.3 <- as.factor(data.node$node.3)
data.node$obs <- seq(1,nrow(data.node))
```


