---
title: "You down with MPT?"
author: "AMS"
date: "1/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#to load data
library(readxl)
library(readr)
#to manipulate data, dplyr, purr, etc and to plot w ggplot2
library(tidyverse)
#Plot tree structure
library(data.tree)
#Frequentist multilevel models
library(lme4)
#Bayesian multilevel models
library(brms)
#pretty plots
library(sjPlot)
```
#### Load data and filer out practice for single participant
```{r}
#load data
library(readxl)
Practice_xls_withpractice <- read_excel("Participant2.xlsx")
Practice_xls <- read_excel("Participant2.xlsx") %>% 
  filter(condition != "practice")
```

#### Batch load all data files from a given working directory:
```{r}
my_files = list.files("~/GitHub/DR2/Specific aim 2/Academy of Aphasia/data")
#Further arguments to read.csv can be passed in ...
all_csv <- lapply(my_files,read.csv)

#Set the name of each list element to its
# respective file name. Note full.names = FALSE to
# get only the file names, not the full path.
names(all_csv) <- gsub(".csv","",
                       list.files("~/GitHub/DR2/Specific aim 2/Academy of Aphasia/data",full.names = FALSE),
                       fixed = TRUE)

```

## Intended MPT structure:
```{r, echo=FALSE}
#Create tree structure
TreeStructure = Node$new("Phonological Input: y*tlji1")
UnrelComp = TreeStructure$AddChild("Phonological Competitor (ytlji = 4)")
PhonProcessing = TreeStructure$AddChild("Phonological Processing (y*tlji2)")
TreeStructure$AddChildNode(PhonProcessing)
PhonComp = PhonProcessing$AddChild("Semantic Competitor (ytlji = 3)")
SemProcessing = PhonProcessing$AddChild("Semantic Processing (y*tlji3)")
PhonProcessing$AddChildNode(SemProcessing)
SemComp = SemProcessing$AddChild("Unrelated Competitor (ytlji = 2)")
Target = SemProcessing$AddChild("Target (ytlji = 1)")

#Add Probabilities 


print(TreeStructure, "p")
Do(TreeStructure$leaves, function(node) SetNodeStyle(node, shape = "rectangle"))
plot(TreeStructure)
```
```

## Wrangle data 
```{r}
#Practice_xls_withpractice <- read_excel("Participant2.xlsx")
#df <- read_excel("Participant2.xlsx") %>% 
#  filter(condition != "practice")

df = practice_xls
df2 = df %>%
  rename(aoi1 = RIGHT_IA_1_SAMPLE_COUNT) %>%
  rename(aoi2 = RIGHT_IA_2_SAMPLE_COUNT) %>%
  rename(aoi3 = RIGHT_IA_3_SAMPLE_COUNT) %>%
  rename(aoi4 = RIGHT_IA_4_SAMPLE_COUNT)

#Following schmit et al, change all values below 5ms to 0 and above to 10 (pg 14, table expalnation.
df2$aoi1 = ifelse(df2$aoi1 < 5, 10, 0)
df2$aoi2 = ifelse(df2$aoi2 < 5, 10, 0)
df2$aoi3 = ifelse(df2$aoi3 < 5, 10, 0)
df2$aoi4 = ifelse(df2$aoi4  < 5, 10, 0)

#create single column showing what aoi participant was looking at
df2$aoi_all = ifelse(df2$aoi1 > 0, 1,
                     ifelse(df2$aoi2 > 0, 2, 
                            ifelse(df2$aoi3 > 0, 3,
                                   ifelse(df2$aoi4 > 0,4, 0))))

df2$y_5 = ifelse(df2$aoi_all == df2$unrelated_location, 2,
                 ifelse(df2$aoi_all == df2$phonemic_location, 4,
                        ifelse(df2$aoi_all == df2$semantic_location, 3,
                               ifelse(df2$aoi_all == df2$target_location, 1, 0)))) 
# yy.list is the hierachical structure of available steps in the tree model. 

library(tidyr)

df2 = df2 %>%
  mutate(y =  na_if(y_5, 0)) %>% 
  fill(y, .direction = "up")

df3 = data.frame(  
  trial = df2$TRIAL_INDEX,
  person = rep(2, nrow(df2)),
  time = df2$BIN_START_TIME,
  item = df2$targetword,
  y = df2$y
)
```

## Triplicate data for number of nodes = 3
```{r}
number.of.nodes <- 3
for (node in 1:number.of.nodes){
  data.copy <- df3
  data.copy$node <- rep(node, nrow(df3))
  if (node==1){data.node <- data.copy} else {data.node <- rbind(data.node,data.copy)}
}
#Sanity check make sure we did the missing data transformation correct
#does step 1 result in missing data in the form of numerical value 0? (yes)
unique(df2$aoi_all)
#Does step 2 result in missing data in the form of numerical value 0? (yes)
unique(df2$y_5)
#does step 3 remove na in the form of zero and replace with nedxt highest number in teh sequence? (yes)
NA_check = data.frame(df2$y_5, df3$y)
```

## Create variable yy, simplified code. 
```{r}
data.node$yy = ifelse(data.node$node== 1 & data.node$y < 4, 1,  #node 1 and correct (sem, unrelated, correct) 
                      ifelse(data.node$node== 1 & data.node$y == 4, 0, #node 1 and phon distractor
                             ifelse(data.node$node== 2 & data.node$y == 4, NA, #node 2 and phonological comp
                             ifelse(data.node$node== 2 & data.node$y == 3, 0, #node 2 and phonological comp
                             ifelse(data.node$node== 2 & data.node$y < 3, 1,  #node 2 and unrelated or correct
                             ifelse(data.node$node== 3 & data.node$y > 2, NA, #node 3 and y = phonological or semantic competitor
                             ifelse(data.node$node== 3 & data.node$y == 2, 0, 1 #node 3 and unrelated competitor, if not unrelated then correct
                             )))))))
unique(data.node$yy)

```

## Remove outcome values thare are not possible at node locations
```{r}
if (length(which(data.node$node==2 & data.node$y==4))>0){
  data.node <- data.node[-which(data.node$node==2 & data.node$y==4),]
}
#Remove y = 3 or 4 for node 3
if (length(which(data.node$node==3 & data.node$y>2.1))>0){
  data.node <- data.node[-which(data.node$node==3 & data.node$y>2.1),]
}

#Test, node 2 has no y = 4, true
test_node_2 = data.node %>%
  filter(node == 2) %>%
  filter(y ==4)
nrow(test_node_2)
#Test, node 3 has no y = 3 or 4, true
test_node_3 = data.node %>%
  filter(node == 3) %>%
  filter(y > 2.1)
nrow(test_node_3)
```

#create fail safe data.node, and data.node.orginal, modify data.node below
```{r}
data.node.original <- data.node <- data.node[order(data.node$node, data.node$person, data.node$trial, data.node$time),]
data.node <- data.node[order(data.node$node, data.node$person, data.node$trial, data.node$time),]

#create variables to use later
data.node$time.coded <-  data.node$time.coded.centered <- rep(0,nrow(data.node))

```

## Create variables node.1, node.2, and node.3
- node.1: if the participant is looking at phonological comp 0, anything else 1.
- node.2 At node 2, if the participant is looking at semantic 0, anything else (note phonological no longer in dataset for node = 2)
-node.3 At node 3, if the participant is looking at unrelated 0, and correct = 1 (note phonological and semantic competitors no longer in dataset for node = 2)

```{r}
#At node 1, if the participant is looking at phonological comeptitor 0 anything else down the tree, 1
data.node$node.1 = ifelse(data.node$y == 4, 0, 1)
#At node 2, if the participant is looking at semantic competitor or phonological, 0 anything else 1
data.node$node.2 = ifelse(data.node$y > 3, 0, 1)
#At node 3, if the participant is looking at correct, 1 anything else 0
data.node$node.3 = ifelse(data.node$y == 2, 0, 1)
```

## Create variable t and c
-t: Whether the person is looking at the target. t=1 if y=1, and t=0 otherwise.
-c: Whether the person is looking at the competitor. c=1 if y=2, and c=0 otherwise.
```{r}
node1_df = data.node %>%
  filter(node == 1)
node1_df$t = ifelse(node1_df$y != 4, 1,0)
node1_df$c = ifelse(node1_df$y !=1,1,0)

node2_df = data.node %>%
  filter(node == 2)
node2_df$t = ifelse(node2_df$y == !3, 1,0)
node2_df$c = ifelse(node2_df$y !=1,1,0)

node3_df = data.node %>%
  filter(node == 3)
node3_df$t = ifelse(node3_df$y == 1, 1,0)
node3_df$c = ifelse(node3_df$y !=1, 1,0)

data.node = rbind(node1_df, node2_df, node3_df)

#SANITY CHECK: ARE THERE NOW 3 NODES AGAIN?
unique(data.node$node)
```

#Create lag variables: t.lag, c.lag, yy.lag
- t.lag: The value of t at the previous time point. Undefined for the first time point.
- c.lag: The value of c at the previous time point. Undefined for the first time point.
- yy.lag: The value of yy at the previous time point. Undefined for the first time point.

```{r}
data.node = data.node %>%
  arrange(node) %>%
  group_by(node, person, item) %>%
  mutate(t.lag = lag(t)) %>%
  mutate(c.lag = lag(c)) %>%
  mutate(yy.lag = lag(yy))%>%
  group_by() %>%
  mutate(time.coded.center = scale(time)) %>%
  filter(!is.na(t.lag))

#SANITY CHECK: Check to make sure t.lag, c.lag, nor yy.lag include NA 
summary(data.node)
```

## Declare variable structure for analyses:
```{r}
#create correct variable structures 
data.node$yy.lag <- as.numeric(data.node$yy.lag)
data.node$t.lag <- as.numeric(data.node$t.lag)
data.node$c.lag <- as.numeric(data.node$c.lag)
data.node$time.coded <- as.numeric(data.node$time.coded)
data.node$time.coded.centered <- as.numeric(data.node$time.coded.centered)
data.node$item <- as.factor(data.node$item)
data.node$person <- as.factor(data.node$person)
data.node$trial <- as.factor(data.node$trial)
data.node$node <- as.factor(data.node$node)
data.node$node.1 <- as.factor(data.node$node.1)
data.node$node.2 <- as.factor(data.node$node.2)
data.node$node.3 <- as.factor(data.node$node.3)
data.node$obs <- seq(1,nrow(data.node))
```

## Load covariates:
```{r}
covariates = read_excel("C:/Users/Brahma/Downloads/VWP_PreTx_assessments_CAC.xlsx")
```

# MPT modeling
## Model 1:
Fixed effects
- node for the intercept γr
- t.lag:node for the fixed lag effect λ_{Tr}
- c.lag:node for the fixed lag effect λ_{Cr}
- PhonologicalVariable:node for the first experimental condition γ_{1r}
- Semantic Variable::node for the first experimental condition γ_{2r}

Random effects
(-1 + node | trial) for the trial random effect δ_{ljir}
(-1 + t.lag:node + c.lag:node + node | person) for random person effects λ_{Tr}, λ_{Cr}, and θ_{jr}
(-1 + t.lag:node + c.lag:node + node | item) for random item effects λ_{Tr}, λ_{Cr}, and β_{ir}

```{r}
## Does not converge: Model.1 <- glmer(yy ~ -1 + node + t.lag:node + c.lag:node + time.coded:node  
##+ (-1 + node |trial), 

# add in interaction for sem and phon: + phon:node +sem:node
Model.1 <- glmer(yy ~ -1 + node + t.lag:node + c.lag:node + time.coded.centered:node 
+ (-1 |trial), 
family = binomial, 
data = data.node)

summary(Model.1)
```