---
title: "You down with MPT?"
author: "AMS"
date: "3/1/2022"
output: html_document
---

```{r, warning=FALSE, message=FALSE, include=FALSE}
#working directory package
library(here)
#to load data
library(readxl)
library(readr)
#to manipulate data, dplyr, purr, etc and to plot w ggplot2
library(tidyverse)
#list manipulation
library(rlist)
#Plot tree structure
library(data.tree)
#Frequentist multilevel models
library(lme4)
#Bayesian multilevel models
library(brms)
#pretty plots
library(sjPlot)
library(httr)
library(stargazer)
```

```{r, message=FALSE, message=FALSE, include=FALSE}
#load data
data = read_csv(here("formatted data", "mpt_analysis_data.csv"))
ivs = read_csv(here("dr2_centered_IVs.csv"))
```

```{r}
# create correct variable structures 
data$yy.lag <- as.numeric(data$yy.lag)
data$t.lag <- as.numeric(data$t.lag)
data$c.lag <- as.numeric(data$c.lag)
data$item <- as.factor(data$item)
data$person <- as.factor(data$id)
data$trial <- as.factor(data$TRIAL_INDEX)
data$node <- as.factor(data$node)
data$node.1 <- as.factor(data$node.1)
data$node.2 <- as.factor(data$node.2)
data$node.3 <- as.factor(data$node.3)
data$obs <- seq(1,nrow(data))

data = data %>%
  group_by(person, item) %>%
  mutate(time = row_number())


```


```{r}
data = data %>%
  group_by(person, item) %>%
  mutate(trial_re = row_number()) %>%
  ungroup() %>%
  mutate(obs = seq(1, nrow(data), 1))
```

### Model fitting, starting small. Unidimensional model with fixed node.

$$ n_{tljir} = \gamma_1 + y^{*'}_{(t-1)ljir^\gamma} + time^{'}_{tljir^\zeta} + \delta_{lji} + \theta_j + \beta_i $$ $$ \gamma_1 = intercept \\
 time^{'}_{tljir^\zeta} = \: fixed \: trend \: effect \: (time_{(1,2...n)}) \\
  y^{*'}_{(t-1)ljir^\gamma} = lag_{t-1}: target: 1, other: 0 \\
 \delta_{lji} = trial\:level \: random \: effect \\
 \theta_j = random \: person \: effect \\
 \theta_j = random \: item \: effect
 $$

\#\#\# loading model output

```{r}
data_uni_node1 = data %>% 
  filter(node == 1)
# 
# Node1mod = brm(yy ~ 1 + time + t.lag +  (1|item) + (1|person) + (1|trial),
#                data = data_uni_node1,
#                prior = c(prior(normal(0, 2.5), class = Intercept),
#                          prior(normal(0, 10), class = b)), #prior on betas
#                  chains = 4,
#                  cores = 4,
#                  family = binomial)


#save(Node1mod, file = "fit1.rda")

node1mod = load("fit1.rda")

```

#Assess model convergence
```{r}
plot(Node1mod)
```

#Posterior predictive check Node1model

```{r}
pp_node1m1 = pp_check(Node1mod, ndraws = 1000)
pp_node1m1
```

```{r}
summary(Node1mod)

mod_sum = summary(Node1mod)
mod_fixed =mod_sum$fixed
mod_ran = rbind(mod_sum$random$item, mod_sum$random$person)
row.names(mod_ran) = c("RE: Intercept", "RE:Person")

mod_out = rbind(mod_fixed, mod_ran)

stargazer(mod_out, type = 'text', summary = FALSE, title = "Table 1: Node 1 with IV t.lag")

```



```{r}
data_uni_node2 = data %>% 
  filter(node == 2)

# Node2mod = brm(yy ~ 1 + time + t.lag +  (1|item) + (1|person) + (1|trial),
#                data = data_uni_node2,
#                prior = c(prior(normal(0, 2.5), class = Intercept),
#                          prior(normal(0, 10), class = b)), #prior on betas
#                  chains = 4,
#                  cores = 4, 
#                  family = binomial)
# save(Node2mod, file = "node2_fit.rda")

load("node2_fit.rda")
```

#Assess model convergence
```{r}
plot(Node2mod)
```

#Posterior predictive check Node1model

```{r}
pp_node2 = pp_check(Node2mod, ndraws = 1000)
pp_node2
```

#Model 2 results:
```{r}
mod_sum = summary(Node2mod)
mod_fixed = mod_sum$fixed
mod_ran = rbind(mod_sum$random$item, mod_sum$random$person)
row.names(mod_ran) = c("RE: Intercept", "RE:Person")
mod_out = rbind(mod_fixed, mod_ran)

stargazer(mod_out, type = 'text', summary = FALSE, title = "Table 1: Node 2 with IV t.lag")
```

```{r}
data_uni_node3 = data %>% 
  filter(node == 3)

# Node3mod = brm(yy ~ 1 + time + t.lag +  (1|item) + (1|person) + (1|trial),
#                data = data_uni_node3,
#                prior = c(prior(normal(0, 2.5), class = Intercept),
#                          prior(normal(0, 10), class = b)), #prior on betas
#                  chains = 4,
#                  cores = 4, 
#                  family = binomial)
# save(Node3mod, file = "node3_fit.rda")
load("node3_fit.rda")
```

#Assess model convergence
```{r}
plot(Node3mod)
```

#Posterior predictive check Node3model

```{r}
pp_node2 = pp_check(Node2mod, ndraws = 1000)
pp_node2
```

#Model 2 results:
```{r}
mod_sum = summary(Node3mod)
mod_fixed = mod_sum$fixed


mod_ran = rbind(mod_sum$random$item, mod_sum$random$person)
row.names(mod_ran) = c("RE: Intercept", "RE:Person")

mod_out = data.frame(rbind(mod_fixed, mod_ran))

stargazer(mod_out, type = 'text', summary = FALSE, title = "Table 3: Node 3 with IV t.lag")

```

```{r}
 # brm(yy ~ 1 + time + t.lag +  (1|item) + (1|person) + (1|trial),
 #               data = data_uni_node1,
 #               prior = c(prior(normal(0, 2.5), class = Intercept),
 #                         prior(normal(0, 10), class = b)), #prior on betas
 #                 chains = 4,
 #                 cores = 4, 
 #                 family = binomial)

load("all_nodes_nodebytime_nodebyt.lag.rda")
```



#Assess model convergence
```{r}
plot(Model4)
```

#Posterior predictive check Node3model

```{r}
pp_check(Model4, ndraws = 1000)

```

```{r}
mod_sum = summary(Model4)
mod_fixed = mod_sum$fixed

mod_ran = rbind(mod_sum$random$item, mod_sum$random$person)
row.names(mod_ran) = c("RE: Intercept", "RE:Person")

mod_out = data.frame(rbind(mod_fixed, mod_ran))

stargazer(mod_out, type = 'text', summary = FALSE, title = "Table 3: Multilevel model with t.lag and c.lag")
```

