---
title: 'Preliminary investigations of visual world paradigm data: Dynamic GLMs and
  Dynamic Tree-Based Item Response Models'
author: "AMS"
date: "12/11/2021"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(dplyr)
library(ggplot2)
library(lme4)
library(readxl)
library(dagitty)
#install.packages('ggdag')
library(ggdag)
```
### Loading Data
When the data is extracted from the edf file it was in a old excel form that neither my computer or R liked.  I opened and saved it as a newer version of an excel workbook, though a csv file would work too.It would be worthwhile to double check to see if this excel file looks correct.

```{r}
library(readxl)
Practice_xls <- read_excel("Output/Practice.xls.xlsx")
```

# Directed Acyclic Graphs
## Need to think about The flow of expected cognitive properties.  Consider Figures below.  Does the expected eye track always match the hypothesized cognitive process order?

Consider the DAG bewlow where, which is the hypothesized input. 
```{r}

Primary_hyp <- dagify(
  Target ~ S,
  S ~ P,
  P ~ Attempt,
  exposure = "Attempt",
  outcome = "Target"
)

tidy_dagitty(Primary_hyp)

ggdag(Primary_hyp, layout = "circle")
```
```{r}
Phon_target<- dagify(
  Target ~ P,
  P ~ Input,
  exposure = "Input",
  outcome = "Target"
)
ggdag(Phon_target, layout = "circle")
``` 
```{r}
sem_target <- dagify(
  S ~ Target ,
  P ~ Input,
  exposure = "Input",
  outcome = "Target"
)

ggdag(sem_target, layout = "circle")
```
```{r}
chaos <- dagify(
  Target ~ P,
  Target ~ S,
  P ~ S,
  S ~ P,
  Target ~ U,
  P ~ U,
  S ~ U,
  U ~ P,
  U ~ S,
  P ~ Input,
  exposure = "Input",
  outcome = "Target"
)
ggdag(chaos, layout = "circle")
``` 

## Data management
First, create dataframe that includes only the variables from from Schmidt et al 2021 and remove original dataframe to save memory.


```{r}
df2 =df %>% 
  filter(RIGHT_INTEREST_AREA_LABEL != ".") 
```
```{r}
df = Practice_xls

df2 = data.frame(  
  trial = df$TRIAL_INDEX,
  person = df$RECORDING_SESSION_LABEL,
  time = df$TIMESTAMP,
  item = df$targetword,
  y_1 = df$RIGHT_INTEREST_AREA_LABEL
)
```
#df contains y_1 as our to be used y variable, need to remove "attention_getter." I reached out to Dr. Cho's doc student to see if ths is what they did in Cho et al., 2021 because the langauge is unclear.  

<center>
"Non-fixation events (blinks and saccades) were treated in the same way, attributing the duration of blink or saccade to the next object that was fixated; as a result, there were no missing polytomous data in the time series for a given trial, prior to modeling multinomial processing."
</center>

```{r}
df3 = df2 %>%
  filter(y_1 != ".") %>%
  filter(y_1 != "ATTENTION_GETTER") 
df4 = df3
df4$y_1 = ifelse(df3$y_1 == "TARGET", 3, 
                      ifelse(df3$y_1 == "SEMANTIC", 2, 
                             ifelse(df3$y_1 == "PHONEMIC", 1, 0
                        )))
data = df4 %>%
  rename(y = y_1)
```



#Merge left and right data sets by create single "fixation index" col. This function will need to be uncommented when we add all participants
```{r}
#data_merge = data
#data_merge$fixation =ifelse(data$RIGHT_FIX_INDEX != ".", data$RIGHT_FIX_INDEX, data$LEFT_FIX_INDEX)
```
  
#Duplicate data for number of nodes. example below n nodes = 2
creates the new data frame with the inclusion of the node variable, first by copying the data for each node and then combining the copies into a single data frame. Note that the number of rows for ```data.node [nrow(data.node)]``` should be equal to $ number.of.nodes×nrow(data)$
```{r}
number.of.nodes <- 2 # change to 4 later
for (node in 1:number.of.nodes){
data.copy <- data
data.copy$node <- rep(node, nrow(data))
if (node==1){data.node <- data.copy} else {data.node <- rbind(data.node,data.copy)}
}
```

#Create variable yy, not exactly sure where this comes from in the paper, but i think its the response to node of interest
```{r}
yy.list <- list(list(1,1,0),list(1,0,NA))
```
# I think that the periods are messing up the code block below within variable ```y```


```{r}
for (observation in 1:nrow(data.node)){
data.node$yy[observation] <- yy.list[[data.node$node[observation]]][[data.node$y[observation]]]
}
```

```{r}
data.node$y = ifelse()
```


#Calculate empircal logit
First, create data.el, as done by naveirmd.github.io...

```{r}
data.el = data.frame(
  trial = data.node$trial,
  item = data.node$item,
  person = data.node$person,
  time = data.node$time,
  node = data.node$node,
  yy = data.node$yy
)
head(data.el)
```

#Assign unique values for trial, time, and node.

```number.variables``` is the number of variables used to estimate the $empirical.logit$

$Empirical.Logit$ is a matrix with each row having unique combinations of $trial$, $time$, and $node$ 

```{r}
unique.trial <- sort(unique(data.el$trial))
unique.time <- sort(unique(data.el$time))
unique.node <- sort(unique(data.el$node))

number.variables <- 3

Empirical.Logit <- matrix(nrow=prod(length(unique.trial),length(unique.time),
length(unique.node)), ncol = (number.variables + 1))
```

#Estimate Empirical Logits
Line 1 in the code above simply sets the initial value for the $row.index$ we’ll use to specify what row of $Empirical.Logit$ we’re currently using. Lines 3-5 indicate that we’re repeating the following process for each unique combination of $trial$, $time$, and $node$. Line 7 extracts the rows of $data.el$ that match the unique combination of $trial$, $node$, and $time$ we’re calculating the empirical logit for, calling this variable $data.el.trial.time.node.$ Line 9 extracts the values for $yy$ from $data.el.trial.time.node$, which will be used to calculate the empirical logit. Lines 10-11 calculate the empirical logit, using the formula ```empirical.logit=log(proportion1−proportion)```, where the proportion is the proportion of $yy.trial.time.node$ equal to one $(∑yylength(yy))$. Lines 17-18 turn Empirical.Logit into a dataframe, with the appropriate names for its variables. Lines 19-20 deal with any cases where $empirical.logit=±∞$, which can occur when proportion=0 or proportion=1. Because the autocorrelations and partial autocorrelations cannot be calculated with infinite empirical logits, we set these values to extremely large values in magnitude (in this case, ±106).

Looking at Empirical.Logit, we should now have the calculated values for empirical logit:
```{r}
row.index <- 1

for (trial in 1:length(unique.trial)){
for (time in 1:length(unique.time)){
for (node in 1:length(unique.node)){

data.el.trial.time.node <- data.el[(data.el$trial==unique.trial[trial] 
& data.el$time==unique.time[time] & data.el$node==unique.node[node]),]
yy.trial.time.node <- data.el.trial.time.node$yy
proportion <- sum(yy.trial.time.node[!is.na(yy.trial.time.node)])/length(yy.trial.time.node)
empirical.logit <- log(proportion/(1 - proportion))
Empirical.Logit[row.index,] <- c(unique.trial[trial], unique.time[time], 
unique.node[node], empirical.logit)
row.index <- row.index + 1
}}}

Empirical.Logit <- data.frame(Empirical.Logit)
names(Empirical.Logit) <- c("trial", "time", "node", "empirical.logit")
Empirical.Logit$empirical.logit[Empirical.Logit$empirical.logit < -10^6] <- -10^6
Empirical.Logit$empirical.logit[Empirical.Logit$empirical.logit > 10^6] <- 10^6
```
#Calculate autocorrelation and Partial Autocorrelations

```{r}
row.index <- 1

for (trial in 1:length(unique.trial)){
for (node in 1:length(unique.node)){

Empirical.Logit.trial.node <- Empirical.Logit[(Empirical.Logit$trial==unique.trial[trial] 
& Empirical.Logit$node==unique.node[node]),]
autocorrelations <- acf(Empirical.Logit.trial.node$empirical.logit, 
lag.max = time.lag.max, plot=FALSE)
autocorrelations <- autocorrelations$acf[2:(time.lag.max + 1)]
partial.autocorrelations <- pacf(Empirical.Logit.trial.node$empirical.logit, 
lag.max = time.lag.max, na.action=na.pass, plot=FALSE)
partial.autocorrelations <- c(partial.autocorrelations$acf)
for (time.lag in 1:time.lag.max){
AC.PAC[(row.index + time.lag - 1),] <- c(unique.trial[trial], unique.node[node], time.lag, 
autocorrelations[time.lag], partial.autocorrelations[time.lag])
}
row.index <- row.index + time.lag.max
}}

AC.PAC <- data.frame(AC.PAC)
names(AC.PAC) <- c("trial", "node", "time.lag", "autocorrelation", "partial.autocorrelation")
```
