---
title: 'Preliminary investigations of visual world paradigm data: Dynamic GLMs and
  Dynamic Tree-Based Item Response Models'
author: "AMS"
date: "12/11/2021"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(dplyr)
library(ggplot2)
library(lme4)
library(readxl)
library(dagitty)
#install.packages('ggdag')
library(ggdag)
```
### Loading Data
When the data is extracted from the edf file it was in a old excel form that neither my computer or R liked.  I opened and saved it as a newer version of an excel workbook, though a csv file would work too.It would be worthwhile to double check to see if this excel file looks correct.

```{r}
#change path to your datafile path
# T12721_1 <- read_excel("C:/Users/alexa/Dropbox/VWP_PRACTICE_REDEPLOY_deploy/results/T12721_1/T12721_1.xlsx")


```

# Directed Acyclic Graphs
## Need to think about The flow of expected cognitive properties.  Consider Figures below.  Does the expected eye track always match the hypothesized cognitive process order?

Consider the DAG bewlow where, which is the hypothesized input. 
```{r}

Primary_hyp <- dagify(
  Target ~ S,
  S ~ P,
  P ~ Input,
  exposure = "Input",
  outcome = "Target"
)

tidy_dagitty(Primary_hyp)

ggdag(Primary_hyp, layout = "circle")
```
```{r}
Phon_target<- dagify(
  Target ~ P,
  P ~ Input,
  exposure = "Input",
  outcome = "Target"
)
ggdag(Phon_target, layout = "circle")
``` 
```{r}
sem_target <- dagify(
  S ~ Target ,
  P ~ Input,
  exposure = "Input",
  outcome = "Target"
)

ggdag(sem_target, layout = "circle")
```
```{r}
chaos <- dagify(
  Target ~ P,
  Target ~ S,
  P ~ S,
  S ~ P,
  Target ~ U,
  P ~ U,
  S ~ U,
  U ~ P,
  U ~ S
  P ~ Input,
  exposure = "Input",
  outcome = "Target"
)
ggdag(chaos, layout = "circle")
``` 

## Data management
First, create dataframe that includes only the variables from from Schmidt et al 2021 and remove original dataframe to save memory.



```{r}
df = Practice_xls

data = data.frame(  
  trial = df$TRIAL_INDEX,
  person = df$RECORDING_SESSION_LABEL,
  time = df$TIMESTAMP,
  item = df$targetword,
  y = df$RIGHT_FIX_INDEX
)
  #Target_fixation = rbinom(nrow(df), size = 1, prob = 0.8),
  #Phon_competitor = rbinom(nrow(df), size = 1, prob = 0.6),
  #Sem_competitor = rbinom(nrow(df), size = 1, prob = 0.4),
  #Unrel_competitor = rbinom(nrow(df), size = 1, prob = 0.1)) #%>%
  #yjti = ACCURACY,
  #AR_phon2sem = ,
  #AR_phon2target = ,
  #AR_sem2target = )
```
  
#Duplicate data for number of nodes. example below n nodes = 2
creates the new data frame with the inclusion of the node variable, first by copying the data for each node and then combining the copies into a single data frame. Note that the number of rows for ```data.node [nrow(data.node)]``` should be equal to $ number.of.nodes×nrow(data)$
```{r}
number.of.nodes <- 2 # change to 4 later
for (node in 1:number.of.nodes){
data.copy <- data
data.copy$node <- rep(node, nrow(data))
if (node==1){data.node <- data.copy} else {data.node <- rbind(data.node,data.copy)}
}
```

#Create variable yy, not exactly sure where this comes from in the paper, but i think its the response to node of interest
```{r}
yy.list <- list(list(1,1,0),list(1,0,NA))
data.node$yy = NA
```
# I think that the periods are messing up the code block below within variable ```y```


```{r}
for (observation in 1:nrow(data.node)){
data.node$yy[observation] <- yy.list[[data.node$node[observation]]][[data.node$y[observation]]]
}
```

```{r}
data.node$y = ifelse()
```


#Calculate empircal logit

```{r}
data.el = data.frame(
  trial = data.node$trial,
  person = data.node$person,
  time = data.node$time,
  node = data.node$node,
  yy = 
)
head(data.el)
```
